Progress Report

### 2023-10-02 Docs/ProgressReport.Rmd
try the debias step

1. Interpolate the eigenfunctions with B-spline basis back to the original grid
2. Debias the eigenvalues with GLMM (mgcv::bam, mgcm::gamm or gamm4:gamm) to be used for Laplace approximation
- we do not need to debias individual scores because the scores for training sample will not be used for out-of-sample prediction

First I'd like to try to fit the debias GLMM model on the full training set with un-interpolated eigenfunctions. The model failed because of vector memory exhaustion. Shall I: 
- Decrease the density of grid (increase bin width)
- Increase memory (re-set to 100GB). But still took a lot of time (more than one hour). The model actually did not finish fitting.

The bias was caused by the misspecification of latent process (assume constant effect over smooth function). And what needs re-evaluation are eigenvalues. Basis functions and mean functions are not biased.

```{r, eval=FALSE}
df_phi <- data.frame(bin=mid, fpca_mod$efunctions[, 1:4])
colnames(df_phi) <- c("bin", paste0("phi", 1:4))

@@

```{r, fig.height=5, fig.width=5}
# mean
interp_mean <- interpSpline(mid, fpca_mod$mu)
full_mean <- predict(interp_mean, x=1:J)
plot(mid, fpca_mod$mu, ylab = "mean", pch = 20)
lines(full_mean$x, full_mean$y, col = "blue")

@@

@@

$$\begin{aligned}
l(\mathbf{Y_u}|\mathbf{\xi}_u) &= \sum_{t_{m_s}<T_u}l(Y_u^s|\mathbf{\xi}_u) \\

@@

g^{-1}(p_s) = \hat{\eta}(t_{m_s}) &= \hat{\mu}_0(t_{m_s})+\sum_{k=1}^K \xi_{uk}\hat{\phi}(t_{m_s})
\end{aligned}$$

>>>>

1. Interpolate the eigenfunctions with B-spline basis back to the original grid
2. Debias the eigenvalues with GLMM (mgcv::bam, mgcm::gamm or gamm4:gamm) to be used for Laplace approximation

- The bias was caused by the misspecification of latent process (assume constant effect over smooth function). 
- What needs re-evaluation are eigenvalues (variance estimates). They will be used in Laplace Approximation for out-of-sample prediction. 
- We do not need to debias individual scores because the scores for training sample will not be used for out-of-sample prediction
- How are the eigenvalues biased? 
- It looks like the FPCA step now estimates mean and eigenfunctions, while each eigenfunction serves as a random slope. Can we fit such model with an offset of mean/intercept function?

First I'd like to try to fit the debias GLMM model on the full training set with un-interpolated eigenfunctions. The model failed because of vector memory exhaustion. Shall I: 
- Decrease the density of grid (increase bin width)?
- Decrease sample size? 
- Increase memory (re-set to 100GB). But still took a lot of time (more than one hour). The model actually did not finish fitting.

```{r, eval=FALSE, class.source = "fold-show"}
df_phi <- data.frame(bin=mid, fpca_mod$efunctions[, 1:4])
colnames(df_phi) <- c("bin", paste0("phi", 1:4))

@@

```{r, fig.height=5, fig.width=5, class.source = "fold-show"}
# mean
interp_mean <- interpSpline(mid, fpca_mod$mu)
full_mean <- predict(interp_mean, x=1:J)
plot(mid, fpca_mod$mu, ylab = "mean", pch = 20)
lines(full_mean$x, full_mean$y, col = "blue")

@@

Perhaps I should try with the simulation data.

@@

\begin{aligned}
l(\mathbf{Y_u}|\mathbf{\xi}_u) &= \sum_{t_{m_s}<T_u}l(Y_u^s|\mathbf{\xi}_u) \\

@@

g^{-1}(p_s) = \hat{\eta}(t_{m_s}) &= \hat{\mu}_0(t_{m_s})+\sum_{k=1}^K \xi_{uk}\hat{\phi}(t_{m_s})
\end{aligned}
